The hydra and the rooster

### 2020-07-18 latex/chapter-ordinals.tex
redaction intro ordinals

@@

These notions are defined in the module 
\href{../src/html/hydras.Prelude.MoreOrders.html}{Prelude.MoreOrders}
 (for any strict order).

@@

\section{Ordinal numbers}
\index{Maths!Ordinal numbers}

@@

\section{Ordinal notations}
In a proof assistant like \coq{},  it is useful to represent ordinals through some data-type, and make arithmetical operations and comparison effectively implemented  through certified functions.

Of course, since the set of all countable ordinals is not countable, it cannot be represented by a language of finite terms. So, we will study ordinal notations just  for countable segments.

\subsection{A type class for ordinal notations}

Let $<$ be some relation on a given type $A$. In order to build an ordinal  notation 
system out of $(A,<)$, we require $<$ to be a strict, well-founded order, and that 
the reflexive closure of $<$ is a total order on $A$.

\index{Coq!Techniques!Type classes}

Using \coq's standard library and type classes, we propose the following definition.

\vspace{4pt}
\noindent\emph{From Module~\href{../src/html/hydras.Prelude.Ordinal_generic.html}{Prelude.Ordinal\_generic}}

\begin{Coqsrc}
Class OrdinalNotation {A:Type}{lt: relation A}(sto:StrictOrder lt)
      (compare : A -> A -> comparison) :=
  { compare_correct :
      forall alpha beta:A,
        CompareSpec (alpha=beta) (lt alpha beta) (lt beta alpha)
                                 (compare alpha beta);
    wf : well_founded lt}.  
\end{Coqsrc}

>>>>

\index{Maths!Ordinal numbers}

@@

These notions are defined (for any strict order) in the module 
\href{../src/html/hydras.Prelude.MoreOrders.html}{Prelude.MoreOrders}.

@@

\subsection{A class for ordinal notations}

Our definition of ordinal notation is inspired by our use of \coq{}: defining inductive data types, predicates, functions, etc, and proving properties of these functions. 
The following type class describes a well-ordered type, and specifies functions for comparing two ordinals, and checking wether a given  ordinal is $0$, a limit or a successor.

\vspace{4pt}
\noindent\emph{From Library~\href{../src/html/hydras.OrdinalNotations.Definitions.html}{OrdinalNotations.Definitions}}

\begin{Coqsrc}
Class OrdinalNotation {A:Type}{lt: relation A}(sto:StrictOrder lt)
      (compare : A -> A -> comparison)  :=
  {   wf : well_founded lt;
      compare_correct :
         forall alpha beta:A,
          CompareSpec (alpha=beta) (lt alpha beta) (lt beta alpha)
                                   (compare alpha beta);
      ZeroLimitSucc_dec : forall alpha,
        {Least alpha} +
        {Limit alpha} +
        {beta: A | Successor alpha beta}
  }. 
\end{Coqsrc}

%%%% ICI %%%%%

@@
